---
title: 'European repository landscape - visualizations'
author: "Niek Van Wettere (https://orcid.org/0000-0002-9455-368X)"
output:
  html_notebook:
    number_section: yes
    toc: yes
  html_document:
    df_print: paged
    toc: yes
---
<br>
<br>

# Introduction
<br>
In this document, the European repository landscape is examined using different visualizations. I focus on (1) repository types (e.g. institutional, discipline-specific etc.) and (2) usage statistics pertaining to repositories.


```{r, message = FALSE}
# load packages
.libPaths("C:/R/library") # load R packages: the path has to be adjusted to where the packages are saved

library(readxl, quietly = T); library(dplyr, quietly = T); library(ggplot2, quietly = T); library(stringr, quietly = T)
```
<br>
The raw data is collected from re3data and Datacite. See below for the type of content.

```{r}

# read data from different source files
data_re3data<-read_excel("Raw_data/20220114_output_re3data_2.xlsx") # from re3data
head(data_re3data)

data_DataCite_repo_names<-read_excel("Raw_data/20220112_output_DataCite_clients.xlsx") # from DataCite
head(data_DataCite_repo_names)

data_DataCite_usage<-read_excel("Raw_data/20220112_output_DataCite_depo_per_repo_2021.xlsx") # from DataCite
colnames(data_DataCite_usage)[1] <- "repo_symbol"
head(data_DataCite_usage)

```
<br>


# Re3data: institutional vs. discipline-specific repositories
<br>
First, the number of repositories is plotted for the countries with at least ten repositories, based on information from re3data (2022/01/14). In this plot, one repository can belong to two or more countries.

```{r}
# plot_1

repo_per_country <- as.data.frame(table(data_re3data$country))
restricted_country_list<-repo_per_country[repo_per_country$Freq > 10, ]$Var1

data_re3data_restricted<-filter(data_re3data, as.character(country) %in% as.character(restricted_country_list))

table_re3data<-as.data.frame(table(as.character(data_re3data_restricted$country), as.character(data_re3data_restricted$type)))
colnames(table_re3data)<-c("country", "type", "Freq")
table_re3data<-filter(table_re3data, type != "")

(repo_plot<-ggplot(table_re3data, aes(x=country, y=Freq, fill = type)) + geom_bar(stat = "identity") + theme_minimal() + coord_flip() + ggtitle("Number of repositories per country") + scale_fill_manual(values = c("disciplinary" = "#d53e4f", "disciplinary+institutional" = "#fc8d59", "institutional" = "#fee08b", "other" = "#762a83", "other+disciplinary" = "#8c510a", "other+disciplinary+institutional" = "#99d594", "other+institutional" = "#3288bd"))) 

```


<br>

Now, the same type of graph is plotted, but this time, the relevant disciplines are also taken into account. 

```{r fig.height=20, fig.width=15}
# plot_2

repo_per_country <- as.data.frame(table(data_re3data$country))
restricted_country_list<-repo_per_country[repo_per_country$Freq > 50, ]$Var1

data_re3data_restricted<-filter(data_re3data, as.character(country) %in% as.character(restricted_country_list))

for (i in (1:nrow(data_re3data_restricted))) {  # eliminate doubles in "subject first" column

  temporary<-str_split(data_re3data_restricted$subject_first[i], "\\+ ?") 
  temporary_unique <- unique(temporary[[1]])
  data_re3data_restricted$subject_first[i]<-paste(temporary_unique, collapse="+")
}

table_re3data_2<-as.data.frame(table(as.character(data_re3data_restricted$subject_first), as.character(data_re3data_restricted$country), as.character(data_re3data_restricted$type)))
colnames(table_re3data_2)<-c("research_discipline", "country", "type", "Freq")
table_re3data_2<-filter(table_re3data_2, type != "" & research_discipline != "NA" & Freq != 0 & research_discipline != "")

(repo_plo_2<-ggplot(table_re3data_2, aes(x=research_discipline, y=Freq, fill = type)) + geom_bar(stat = "identity") + theme_minimal() + coord_flip() + ggtitle("Number of repositories per country and discipline") + facet_wrap(~ country, ncol = 1)) + scale_fill_manual(values = c("disciplinary" = "#d53e4f", "disciplinary+institutional" = "#fc8d59", "institutional" = "#fee08b", "other" = "#762a83", "other+disciplinary" = "#8c510a", "other+disciplinary+institutional" = "#99d594", "other+institutional" = "#3288bd")) 

```
<br>
Since certain repositories occur more than once in the data, the plot below aims to ajust for this by joining multiple countries in these cases. Moreover, only repository types with a frequency higher than 5 are retained in this plot.

```{r fig.height=13, fig.width=10}

# some repositories occur more than once in the data

unique_ids <- unique(data_re3data$ID)
diff_repo <- length(data_re3data$ID) - length(unique_ids) # difference between total number of repo and total number of unique repo

doubles <- list()  # locate all the doubles in the data
doubles_without_first <- list()
for (i in unique_ids) {
  loc <- which(data_re3data$ID == i)
  
  if (length(loc) >= 2) {
    doubles<-append(doubles, list(loc))
    doubles_without_first <- append(doubles_without_first, list(loc[2:length(loc)]))
  }}

re3data_limited<-data_re3data
re3data_limited$country <- as.character(re3data_limited$country)
for (j in 1:length(doubles)) {
  countries <- paste(as.character(data_re3data[doubles[[j]],]$country), collapse = "+")  # merge countries for same repository
  
  re3data_limited[doubles[[j]][1],4] <- countries
  }

to_be_removed <- unlist(doubles_without_first)
re3data_limited <- re3data_limited[-to_be_removed, ] 

disc_instit_data<-as.data.frame(table(as.character(re3data_limited$country), as.character(re3data_limited$type)))
disc_instit_data_2<-filter(disc_instit_data, Freq != 0 & Var2 != "")
colnames(disc_instit_data_2)<-c("country", "type", "Freq")

disc_instit_data_3<-filter(disc_instit_data_2, Freq >= 5) # select only most frequent ones to plot
ggplot(disc_instit_data_3, aes(x=country, y=Freq, color = type)) + geom_point(size = 5) + theme_minimal() + coord_flip() + ggtitle("Repository frequency per country and repository type") + scale_color_manual(values = c("disciplinary" = "#d53e4f", "disciplinary+institutional" = "#fc8d59", "institutional" = "#fee08b", "other" = "#762a83", "other+disciplinary" = "#8c510a"))

```



# Repositories according to usage statistics
<br>
Secondly, European research data repositories are ranked according to how much they were used in publication year 2021, based on metadata from DataCite. The first plot below ranks the repositories taking into account all types of submissions. 


```{r fig.height=10, fig.width=10, warning = FALSE}

DataCite_new <- merge(data_DataCite_usage, data_DataCite_repo_names, by = "repo_symbol", all = TRUE)
DataCite_new_2 <- DataCite_new[order(-DataCite_new$total_depo_2021),]
DataCite_new_3 <- DataCite_new_2[1:50,]

require(scales)
ggplot(DataCite_new_3, aes(x=reorder(repo_name, total_depo_2021), y=total_depo_2021)) + geom_point(size = 5) + theme_minimal() + coord_flip() + ggtitle("Number of submitted records for top-50 repositories") + xlab("Repositories") + ylab("Frequency of submissions in 2021") + scale_y_continuous(labels = scales::comma, limits = c(0,410000)) 
```
<br>

The tabular data underlying the plot above:

```{r}
(DataCite_new_3)
```
<br>
Next, the type of submission is restricted to "dataset" in the plot below.



```{r fig.height=10, fig.width=10, warning = FALSE}
DataCite_new_dataset_count <- filter(DataCite_new, count_datasets != "NA" & count_datasets != "error")
DataCite_new_dataset_count$count_datasets <- as.numeric(DataCite_new_dataset_count$count_datasets)
DataCite_new_dataset_count <- DataCite_new_dataset_count[order(-DataCite_new_dataset_count$count_datasets),]
DataCite_new_dataset_count_2 <- DataCite_new_dataset_count[1:50,]

require(scales)
ggplot(DataCite_new_dataset_count_2, aes(x=reorder(repo_name, count_datasets), y=count_datasets)) + geom_point(size = 5) + theme_minimal() + coord_flip() + ggtitle("Number of submitted datasets for top-50 repositories") + xlab("Repositories") + ylab("Frequency of submissions in 2021") + scale_y_continuous(labels = scales::comma, limits = c(0,410000)) 

```
<br>
The tabular data underlying the plot above:

```{r}
(DataCite_new_dataset_count_2)
```

